---
title: "Dyad-Year Example: A Basic Replication/Extension of Bremer (1992)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r ss-setup, message=F}
library(tidyverse)
library(peacesciencer)
library(lubridate)
library(stevemisc)
library(modelsummary)

packageVersion("peacesciencer")
Sys.Date()
```

[Bremer's (1992)](https://doi.org/10.1177/0022002792036002005) "dangerous dyads" design is not necessarily the seminal analysis of inter-state conflict in the Correlates of War tradition, but it's arguably the seminal template by which one could analyze conflict between states. [The package's manuscript](https://svmiller.com/peacesciencer/ms.pdf) offers a basic extension/replication of this form of analyze, the skeleton of which is ubiquitous in every dyad-year analysis.

```{r}
create_dyadyears(directed = FALSE, subset_years = c(1816:2010)) %>%
  filter_prd() %>%
  # add conflict information from GML-MID data, pipe to next function
  add_gml_mids(keep = NULL) %>%
  add_spells() %>%
  add_nmc() %>%
  add_democracy(keep="aeuds") %>%
  add_atop_alliance() %>% # change from manuscript version
  # likewise, this function below replaces add_sdp_gdp() since version 1.2
  add_sim_gdp_pop(keep="mrgdppc") -> Data

# Some recoding
Data %>%
  mutate(landcontig = ifelse(conttype == 1, 1, 0)) %>%
  mutate(cowmajdyad = ifelse(cowmaj1 == 1 | cowmaj2 == 1, 1, 0)) %>%
  # Create estimate of militarization as milper/tpop
  # Then make a weak-link
  mutate(milit1 = milper1/tpop1,
         milit2 = milper2/tpop2,
         minmilit = ifelse(milit1 > milit2,
                           milit2, milit1)) %>%
  # create CINC proportion (lower over higher)
  mutate(cincprop = ifelse(cinc1 > cinc2,
                           cinc2/cinc1, cinc1/cinc2)) %>%
  # create weak-link specification using Quick UDS data
  mutate(mindemest = ifelse(aeuds1 > aeuds2,
                            aeuds2, aeuds1)) %>%
  # Create "weak-link" measure of jointly advanced economies
  mutate(minmgdppc = ifelse(mrgdppc1 > mrgdppc2,
                             mrgdppc2, mrgdppc1)) -> Data

Data %>%
  r2sd_at(c("cincprop", "mindemest", "minmgdppc", "minmilit")) -> Data

modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + atop_defense +
               mindemest + minmgdppc + minmilit +
               gmlmidspell + I(gmlmidspell^2) + I(gmlmidspell^3), data= Data,
             family=binomial(link="logit"))

```


In table form:

```{r}
modelsummary(list("Model 1" = modDD), 
             title = 'A "Dangerous Dyads" Analysis of Non-Directed Dyad-Years from {peacesciencer}',
             stars = c('*' = .05, '+' = .1), 
             gof_map = c("nobs"),
             coef_map = c("landcontig" = "Land Contiguity", 
                          "cincprop" = "Dyadic CINC Proportion (Lower/Higher)", 
                          "cowmajdyad" = "CoW Major Power in Dyad",
                          "atop_defense" = "Defense Pact", 
                          "mindemest" = "Dyadic Democracy (Weak-Link)", 
                          "minmgdppc" = "Dyadic GDP per Capita (Weak-Link)", 
                          "minmilit" = "Dyadic Militarization (Minimum)",
                          "gmlmidspell" = "t",
                          "I(gmlmidspell^2)" = "t^2",
                          "I(gmlmidspell^3)" = "t^3",
                          "(Intercept)" = "Intercept"),
             align=c("lc")) 
```
