---
title: "State-Year Example: A Basic Replication/Extension of Fearon and Laitin (2003)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r ss-setup, message=F}
library(tidyverse)
library(peacesciencer)
library(lubridate)
library(stevemisc)
library(modelsummary)

packageVersion("peacesciencer")
Sys.Date()
```

[Fearon and Laitin (2003)](https://doi.org/10.1017/S0003055403000534) have one of the most influential state-year analyses of civil war onset to date and their research design serves as a basic template for what such a state-year analysis would resemble. What follows is not a perfect reproduction of their exact design, but an easy extension and replication of it in `{peacesciencer}`. [The package's manuscript](https://svmiller.com/peacesciencer/ms.pdf) offers a bit more detail than I include here. You might also find [Gibler and Miller (2014)](https://doi.org/10.1177/0022343314531003) interesting since some of the package's functionality is derived from what went into that particular extension of Fearon and Laitin (2003).

```{r}
create_stateyears(system = 'gw', subset_years = c(1946:2024)) %>%
  add_ucdp_acd(type= "intrastate", only_wars = FALSE) %>%
  select(-maxintensity, -conflict_ids) %>%
  add_spells() %>%
  add_democracy(keep = "aeuds") %>% 
  # ^ changed from the manuscript version
  add_creg_fractionalization() %>%
  add_sim_gdp_pop(keep = c("mrgdppc", "pwtpop"))  %>%
  # ^ replaced add_sdp_gdp()
  add_rugged_terrain() -> Data

create_stateyears(system = 'gw', subset_years = c(1946:2024)) %>%
  add_ucdp_acd(type= "intrastate", only_wars = TRUE) %>%
  select(-maxintensity, -conflict_ids) %>%
  add_spells(ongo = TRUE, conflict_event_type = "onset") %>%
  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
  left_join(Data, .) -> Data

Data %>%
  arrange(gwcode, year) %>%
  log_at(c("mrgdppc", "pwtpop")) %>%
  lag_at(c("aeuds", "ln_mrgdppc", "ln_pwtpop"),
         .by = "gwcode") -> Data


modCW <- list()
summary(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_ln_mrgdppc + l1_ln_pwtpop  +
                    l1_aeuds + I(l1_aeuds^2) +
                    newlmtnest + ethfrac + relfrac +
                    ucdpspell + I(ucdpspell^2) + I(ucdpspell^3), data=subset(Data),
                  family = binomial(link="logit")))

summary(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_ln_mrgdppc + l1_ln_pwtpop  +
                    l1_aeuds + I(l1_aeuds^2) +
                    newlmtnest + ethfrac + relfrac +
                    war_ucdpspell + I(war_ucdpspell^2) + I(war_ucdpspell^3), data=subset(Data),
                  family = binomial(link="logit")))
```


In table form:

```{r}
modelsummary(modCW, 
             title = 'A Civil Conflict Analysis of Gleditsch-Ward State-Years in {peacesciencer}',
             stars = c('*' = .05, '+' = .1), gof_omit = "IC|F|Log.|R2$",
             coef_map = c("l1_ln_mrgdppc" = "GDP per Capita (Lagged)", 
                          "l1_ln_pwtpop" = "Population Size (Lagged)", 
                          "l1_aeuds" = "Extended UDS (Lagged)",
                          "I(l1_aeuds^2)" = "Extended UDS^2 (Lagged)", 
                          "newlmtnest" = "% Mountainous Terrain (Logged)", 
                          "ethfrac" = "Ethnic Fractionalization", 
                          "relfrac" = "Religious Fractionalization",
                          "ucdpspell" = "t",
                          "I(ucdpspell^2)" = "t^2",
                          "I(ucdpspell^3)" = "t^3",
                          "war_ucdpspell" = "t",
                          "I(war_ucdpspell^2)" = "t^2",
                          "I(war_ucdpspell^3)" = "t^3",
                          "(Intercept)" = "Intercept"),
             align=c("lcc"))
```
