---
title: "`{peacesciencer}` Data Versions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F}
library(tidyverse)
library(peacesciencer)
library(kableExtra)
```

These are the data versions available in `{peacesciencer}`. Do note the user can find specific data versions with more targeted use of the `ps_version()` function in this package. Without an argument, `ps_version()` produces a data frame of all the data versions in this package. 


```{r}
ps_version() %>%
  kbl(., caption = "Data Versions in `{peacesciencer}`",
      align =c("c", "l", "c", "c"))  %>%
  kable_styling(position = "center", full_width = F, bootstrap_options = "striped")
```

Users can use the `bibtexkey` information to search [the dataframe of citations](http://svmiller.com/peacesciencer/reference/ps_bib.html) in this package. For example, we can use `ps_cite()` to get a full citation for Carter and Smith's (2020) estimates of leader willingness to use force.


```{r}
ps_cite("cartersmith2020fml", column = "bibtexkey")
```
