---
title: "`{peacesciencer}` Data Versions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r vig-versions-setup, message=F}
library(tidyverse)
library(peacesciencer)
library(kableExtra)
```

These are the data versions available in `{peacesciencer}`. Do note the user can find specific data versions with more targeted use of the `ps_version()` function in this package. Without an argument, `ps_version()` produces a data frame of all the data versions in this package. 


```{r}
ps_version() %>%
  kbl(., caption = "Data Versions in `{peacesciencer}`",
      align =c("c", "l", "c", "c"), 
      booktabs = TRUE, longtable = TRUE)  %>%
  kable_styling(position = "center", full_width = F, 
                bootstrap_options = "striped") %>%
  row_spec(0, bold=TRUE)
```

Every data set that is used by a function in this package is included in this table (and with the `ps_version()` function). Users can see a category of the type of data (which can be used for more careful searches of the underlying data), a description of the data set, a version number associated with that data set in `{peacesciencer}`, along with a BibTeX key. Users should interpret data versions as years as instances where the data are not formally versioned, per se, and the year corresponds with a year of a last update or a year of publication. For example, [Anders et al. (2020)](https://doi.org/10.1093/isq/sqaa013) released their state-year simulations of population, surplus domestic product, and gross domestic product in 2020. The data are not formally versioned and the year corresponds with, in this case, the publication.

Users can use the BibTeX key (`bibtexkey`) to search [the dataframe of citations](http://svmiller.com/peacesciencer/reference/ps_bib.html) in this package. For example, we can use `ps_cite()` to get a full citation for Carter and Smith's (2020) estimates of leader willingness to use force.


```{r}
ps_cite("cartersmith2020fml", column = "bibtexkey")
```
